### 2022-03-28 paper/localpsf.tex
reorganizing intro based on Noemi's suggestions

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$) and let $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the following form:
\begin{equation}
	\label{eq:kernel_representation}
	(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy.
\end{equation}
The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.

@@

Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points. The \emph{impulse response} of $\Aop$ associated with the point $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
\begin{equation}
\label{eq:impulse_response_defn}
\impulseresponse_x(y) := \Aker(y, x),
\end{equation} 
which results from keeping $x$ fixed and viewing $\Phi(y,x)$ as a function of $y$. The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
\begin{equation}
\label{eq:impulse_response_delta_action}
\impulseresponse_x = \left( \Aop \delta_x^* \right)^*,
\end{equation}
where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions. This is described in Appendix \ref{app:distributions}.

We assume:
\begin{itemize}
	\item The supports of the impulse responses $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
	\item The integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness. The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
\end{itemize}
Because of these assumptions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop$ to a small number of polynomial functions. We use the support estimates to compute scattered impulse responses in batches (see Figure \ref{fig:batches_intro}). This batching allows us to recover many impulse responses using a small number of operator actions $u \mapsto \Aop u$. The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme.

@@

Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of continuous Galerkin finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$. We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(p_i)=1$ and $\febasis_i(p_j)=0$ if $i\neq j$. 
For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$, i.e.,
\begin{equation}
\label{eq:fem_coeff_basis}
u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
\end{equation}
Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries 
\begin{equation*}
\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,N. 
\end{equation*}
Here $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries
\begin{equation*}
\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,N,
\end{equation*}
the space $\mathbb{R}^N_\mathbf{M}$ is $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$. Direct calculation shows that $\mathbf{R}^N_\mathbf{M}$ and $\mathbf{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.
%, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\rho_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\rho}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\rho_h(u) = \boldsymbol{\rho}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\rho_h, \nu_h \in V_h'$.

>>>>

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$). We seek to approximate integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ that may be written in the following form:
\begin{equation}
	\label{eq:kernel_representation}
	(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy.
\end{equation}
The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which is not easily accessible to us. The linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$. Here $L^2(\Omega)'$ denotes the topological dual of $L^2(\Omega)$ (space of all continuous linear functionals).

@@

%Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points. 
The \emph{impulse response} of $\Aop$ associated with the point $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
\begin{equation}
\label{eq:impulse_response_defn}
\impulseresponse_x(y) := \Aker(y, x),
\end{equation} 
which results from keeping $x$ fixed and viewing $\Phi(y,x)$ as a function of $y$. The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
\begin{equation}
\label{eq:impulse_response_delta_action}
\impulseresponse_x = \left( \Aop \delta_x^* \right)^*,
\end{equation}
where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions. This is described in Appendix \ref{app:distributions}.

We assume (1) the supports of the impulse responses $\impulseresponse_x$ are contained (or approximately contained) in localized regions, and (2) the integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness. The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
Because of these assumptions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop$ to a small number of polynomial functions. We use the support estimates to compute scattered impulse responses in batches (see Figure \ref{fig:batches_intro}). This batching allows us to recover many impulse responses using a small number of operator actions $u \mapsto \Aop u$. The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme.

@@

Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of continuous Galerkin finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$. We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(p_i)=1$ and $\febasis_i(p_j)=0$ if $i\neq j$. 
For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
\begin{equation}
\label{eq:fem_coeff_basis}
u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
\end{equation}
Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries 
\begin{equation*}
\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,N. 
\end{equation*}
Here $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries
\begin{equation*}
\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,N,
\end{equation*}
the space $\mathbb{R}^N_\mathbf{M}$ is $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$. Direct calculation shows that $\mathbf{R}^N_\mathbf{M}$ and $\mathbf{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.
%, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\rho_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\rho}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\rho_h(u) = \boldsymbol{\rho}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\rho_h, \nu_h \in V_h'$.
